Prelim Outline

### 2018-07-17 prelim/outline.tex
Outline cleaned, ready to discuss with lucas

\documentclass{article}
\usepackage{graphicx}
\usepackage{color}
\title{Prelim Outline (ROUGH)}
\author{Shivesh Pathak}

@@

\color{red} Red \color{black} is incomplete preliminary work.
\color{blue} Blue \color{black} is work for the future.

\section{Motivation/Background}
\begin{enumerate}
\item People want to construct an accurate low-energy model Hamiltonian for the cuprates.
\item Hole doped cuprates have been studied for a long time and have many degrees of freedom: O p and Cu d orbitals both play roles.
\item Electron doped cuprates are less studied and we believe they are simpler since only the Cu d orbitals play a major role.
\item We plan on building an low-energy model for the electron doped cuprate SrCuO$_2$ using a density-matrix downfolding method and accurate many-body QMC calculations.
\end{enumerate}

\section{Preliminary SCF calculations}
\subsection{Undoped (x=0)}
\begin{enumerate}
\item We did eight unit cell PBE0 calculations to calculate the energies of states with singly occupied Cu d$_{x^2-y^2}$ orbitals but with different spin textures.

@@

\item The results of this fitting on seven different spin textures yields good results for two different basis sets (BFD Cutoff, BFD PBC 0.2)

\item The high $R^2$ values of these fits indicate that a simple AFM Heisenberg model should be sufficient to capture the low-energy behavior of the undoped cuprates.
\end{enumerate}

\includegraphics[width=0.5\textwidth]{../undoped/PBE0_8/J_fit_naive.pdf}
\includegraphics[width=0.5\textwidth]{../undoped/PBE0_8/J_fit_naive_BFD0p2.pdf}

\subsection{Overdoped (x=0.25)}
\begin{enumerate}
\item We similarly did seventeen x=0.25 doped calculations in PBE0, from we got four low energy states. Here low-energy is defined as states with energy less than 0.3 eV from the lowest energy state we had, a collinear doped state.

\item We then used the band structures and energies of these four low energy states to fit to a naive model with $t, t^\prime, K, J$ parameters: NN-hopping, NNN-hopping, NN-Kondo term, and NN-Heisenberg-AFM term

@@

\item The presence of $t, t^\prime$ parameters were necessary to get the correct curvature of bands near the Fermi level, and the $K, J$ parameters were necessary to get both the spacing between bands (K) and the total energies (K,J).

\item We map each of the PBE0 states into the model Hilbert space by breaking the occupied bands into a) fully occupied and b) partially occupied. The orbitals from a) then build a spin density that defines the variables $\vec{S_i}$ and the orbitals from b) define the variables $c_i^\dagger$.

\item We can therefore create a single body  Hamiltonian where we treat the spins $\vec{S_i}$ as a fixed background upon which the itenerant $c_i^\dagger$ electrons can move. We can solve this model Hamiltonian and fit to the PBE0 band structure, and also compare the total energy of the model to the PBE0 total energy.

\item We use a prior where $J$ is held as the undoped value. This is because in a metallic/superconducting doped state, there are charge fluctuations locally. Therefore we expect that on average the total charge locally between nearest neighbors will be the undoped value, and since the Heisenberg model only has local coupling we don't imagine $J$ should change too much.

\item Since we are fitting to two different quantities (total energy and band structure) we used a Pareto efficiency analysis to get the best balance between a fit to total energy and a fit to band structure. The cost function is shown below, and $w_E$ varies in the plot shown even lower for the two different basis sets from before.

@@

@@

\includegraphics[width=0.5\textwidth]{../doped_fv/PBE0_8/tb_fit/pareto4.pdf}
\linebreak
\color{red} Need to include plot for PBC 0.2 basis set.

@@

\item For the BFD cutoff basis set we get the parameter values $t,t^\prime,K,J=  1.04, 0.53, -0.15, 0.18 eV$

\item The two basis sets do not agree quantitatively, but this isn't too unexpected because it's such a rough fitting method, but qualitatively they do: AFM J, FM K, repulsive NNN hopping, attractive NN hopping.

\includegraphics[width=\textwidth]{../doped_fv/PBE0_8/tb_fit/pareto4_high.pdf}
\color{red} Need to include plot for PBC 0.2 basis set.
\end{enumerate}

\subsection{Optimally doped (x=0.125)}
\color{red} Need to do these calculations, but analysis is identical to above.

\color{black}
\section{Density matrix downfolding (DMD) calculations}
\subsection{Basis}

\begin{enumerate}
\item In order to do our DMD calculations we need a single particle basis to evaluate the density matrix on.  We use intrinsic atomic orbitals (IAOs) as our one particle basis. Justification for this choice is provided below.

\item \textbf{Undoped (x=0)}]
\begin{enumerate}
\item Shown below are the occupations of different IAOs for the checkerboard spin texture for the BFD Cutoff and BFD PBC 0.2 basis sets.

\item Clearly the IAOs show that the primary variation between occupation is in the $3d_{x^2-y^2}$ orbital, namely the spin changes from site to site. The oxygen orbitals do not play a significant role, indicating that this basis captures the qualitative features that we expect from the undoped state.

\item Another important feature is that for each copper atom there are TWO $3d_{x^2-y^2}$ IAOs: one for the occupied spin and one of the unoccupied spin.

\end{enumerate}
\includegraphics[width=0.5\textwidth]{../undoped/PBE0_8/CHK/dens_iao2_cu.pdf}
\linebreak
\color{red} Need to include plot for PBC 0.2 basis set.

\color{black}
\item \textbf{Overdoped (x=0.25)}
\begin{enumerate}
\item For the doped calculation, we did the same IAO analysis as for the x=0 case. We found that again the $3d_{x^2-y^2}$ orbital has the primary variation in occupation.

\item Unlike the x=0 case, once doped the occupation number changes as well as the spin. Below is shown the difference in occupation of the spin-down and spin-up $3d_{x^2-y^2}$ orbitals of the doped collinear state compared to the UNDOPED collinear state.

\item We see that the occupation of the dominant primary channel does not vary, and we can therefore attribute those orbitals to the fixed magnetic moments, and the secondary spin channel does vary and we can attribute those orbitals to the itinerant electrons.

\includegraphics[width=0.5\textwidth]{../doped_fv/PBE0_8/COL0/iao/sub_occd.pdf}
\linebreak
\color{red} Need to include plot for PBC 0.2 basis set.

\color{black}
\item Therefore we see that the basis elements we will be constructing our DM out of are the two different $3d_{x^2-y^2}$ IAOs: the orbital corresponding to the primary spin channel will be used to calculate DM elements of the fixed spin moments $\vec{S_i}$ and the secondary spin channel IAO will be used to calculate DM elements of the itinerant electrons $c_i^\dagger$
\end{enumerate}

\item \textbf{Optimally doped (x=0.125)}
\color{red} Need to do this, should be identical analysis to x=0.25.
\end{enumerate}

\color{blue}
\subsection{States}
\begin{enumerate}
\item \textbf{Undoped (x=0)} For the undoped calculations, we have seven different spin textures. We can just use these for the model fitting since we will be fitting to a single parameter, $J$.

@@

\color{red}
\item \textbf{Optimally doped (x=0.125)}
\begin{enumerate}
\item Need to do this, should be similar.
\end{enumerate}
\color{blue}
\end{enumerate}

>>>>

\documentclass{article}
\usepackage{graphicx}
\usepackage{color}
\title{Prelim Outline}
\author{Shivesh Pathak}

@@

Text in \color{blue} blue \color{black} are things that need to be planned.

\section{Motivation/Background}
\begin{enumerate}
\item We want to construct an accurate low-energy model Hamiltonian for the cuprates.
\item Hole doped cuprates have been studied for a long time, and the relevant degrees of freedom are both O p and Cu d orbitals. 
\item Electron doped cuprates are less studied and we believe they are simpler since the O p orbitals should play a much smaller role.
\item We plan on building an low-energy model for the electron doped cuprate SrCuO$_2$ using a density-matrix downfolding method and accurate many-body QMC calculations.
\end{enumerate}

\section{Preliminary calculations: DFT}
\subsection{Undoped (x=0)}
\begin{enumerate}
\item We did eight unit cell DFT calculations with a PBE0 functional to calculate the energies of states with singly occupied Cu d$_{x^2-y^2}$ orbitals but with different spin textures.

@@

\item The results of this fitting on seven different spin textures yields good results for two different basis sets (BFD Cutoff, BFD PBC 0.2).

\item For every SCF calculation done the BFD PBC 0.2 basis set had higher total energy than the BFD Cutoff, so I will only present results for the Cutoff basis set.

\item The high $R^2$ values of the fit indicates that a simple AFM Heisenberg model should be sufficient to capture the low-energy behavior of undoped SrCuO$_2$.
\end{enumerate}

\includegraphics[width=0.5\textwidth]{../undoped/PBE0_8/J_fit_naive.pdf}

\subsection{Overdoped (x=0.25)}
\begin{enumerate}
\item We similarly did seventeen x=0.25 doped calculations in PBE0, from which we got four low energy states.

\item Similarly to the x=0 calculations, the x=0.25 PBE0 calculations yielded lower total energy for the BFD Cutoff basis. All results shown are for that basis.

\item Low-energy is defined as states with energy less than 0.3 eV from the lowest energy state we had, a collinear doped state.

\item We used the band structures and total energies of these four low energy states to fit to a naive model with $t, t^\prime, K, J$ parameters: NN-hopping, NNN-hopping, NN-Kondo term, and NN-Heisenberg-AFM term

@@

\item We map each of the PBE0 states into the model Hilbert space by breaking the occupied bands into a) fully occupied and b) partially occupied.

\item The orbitals from a) then build a spin density that defines the variables $\vec{S_i}$ and the orbitals from b) define the variables $c_i^\dagger$.

\item We then create a single body  Hamiltonian where we treat the spins $\vec{S_i}$ as a fixed background upon which the itinerant $c_i^\dagger$ electrons can move.

\item We solve this model Hamiltonian and fit to the PBE0 band structure and further compare the total energy of the model to the PBE0 total energy.

\item Since we are fitting to two different quantities we used a Pareto efficiency analysis to get the best balance between a fit to total energy and a fit to band structure.

\item The cost function is shown below, and $w_E$ varies in the plot shown below.

@@

\item In the fitting we hold the value of $J=0.178$ eV fixed as a prior, which is the optimal value from the x=0 fit. \textcolor{red}{Why?}

@@

\includegraphics[width=0.5\textwidth]{../doped_fv/PBE0_8/tb_fit_0p25/pareto4.pdf}
\linebreak

@@

\item The parameter values at this point on the Pareto plot are $t,t^\prime,K,J=  1.04, 0.53, -0.15, 0.178 $eV.

\item Below is a figure with a fit to the total energies and band structure.

\includegraphics[width=0.7\textwidth]{../doped_fv/PBE0_8/tb_fit_0p25/pareto4_low.pdf}
\end{enumerate}

\subsection{(Near) optimally doped (x=0.125)}
\begin{enumerate}
\item The fitting for x=0.125 is identical to x=0.25 except that the states in PBE0 we use are different.

\item As we had for the x=0.25 calculation, we got four low energy states for x=0.125, and all results below are for the BFD Cutoff basis.

\item Below is the Pareto efficiency plot using the four low energy x=0.125 states and using the cost function from before. The blue square is the point we chose as most efficient.

\includegraphics[width=0.5\textwidth]{../doped_fv/PBE0_8/tb_fit_0p125/pareto4.pdf}
\linebreak

\item The parameter values at this point on the Pareto plot are $t,t^\prime,K,J=  1.15,  0.39, -0.29, 0.18 $eV.

\item Below is a figure with a fit to the total energies and band structure.

\includegraphics[width=0.7\textwidth]{../doped_fv/PBE0_8/tb_fit_0p125/pareto4_sel.pdf}
\end{enumerate}

\subsection{Conclusions}
\begin{enumerate}
\item Fitting to our PBE0 results provides some evidence that our low energy model with four terms $t, t^\prime, K, J$ may describe the low energy Hilbert space of SrCuO$_2$ accurately.

\item The table below shows the variation in the model parameters under doping, in eV.

\begin{center}
\begin{tabular}{ c c c c c }
Doping & t & t$^\prime$ & K & J \\ 
\hline
x=0 & 0 & 0 & 0 & 0.178 \\
x=0.125& 1.15&  0.39& -0.29& 0.178 \\
x=0.25 & 1.04& 0.53& -0.15&0.178 \\
\end{tabular}
\end{center}

\item This fitting was preliminary and is not a systematic or rigorous fitting procedure.

\item My thesis project will be based to use the density matrix downfolding procedure to conduct a more rigorous version of model fitting.
\end{enumerate}

\section{Density matrix downfolding (DMD) calculations}
\subsection{Basis}

\begin{enumerate}
\item In order to do our DMD calculations we need a single particle basis to evaluate the density matrix on.  We use intrinsic atomic orbitals (IAOs) as our one particle basis.

\item \textbf{Undoped (x=0)}]
\begin{enumerate}
\item Shown below are the occupations of different IAOs for the checkerboard spin textured PBE0 state.

\item The IAOs show that the primary variation between occupation is in the spin of an unpaired electron in the $3d_{x^2-y^2}$ orbital.

\item The oxygen orbitals do not play a significant role.

\item Another important feature is that for each copper atom there are TWO $3d_{x^2-y^2}$ IAOs: one for the majority spin and one of the minority spin.

\end{enumerate}
\includegraphics[width=0.5\textwidth]{../undoped/PBE0_8/CHK/dens_iao2_cu.pdf}
\linebreak

\item \textbf{Overdoped (x=0.25)}
\begin{enumerate}
\item For the doped calculation, we did the same IAO analysis as for the x=0 case.

\item Unlike the x=0 case, once doped the occupation number changes as well as the spin.

\item Below is shown the difference in occupation of the spin-down and spin-up $3d_{x^2-y^2}$ orbitals of the doped collinear state MINUS to the undoped collinear state.

\item The occupation of the dominant primary channel does not vary but that of the the secondary spin channel does.

\item We can therefore attribute the prior orbitals to to the fixed magnetic moments and the latter to the itinerant electrons.

\includegraphics[width=0.5\textwidth]{../doped_fv/PBE0_8/COL0/iao/sub_occd.pdf}

\item We find that the majority spin IAOs relax in a localized manner with on site $d_{x^2-y^2}$ character, while the minority spin IAOs relax with $O p$ character across the calculation cell.

\includegraphics[width=0.5\textwidth]{../doped_fv/PBE0_8/COL0/iao/DIFF_0_10.pdf} 
\linebreak

\includegraphics[width=0.5\textwidth]{../doped_fv/PBE0_8/COL0/iao/DIFF_0_810.pdf}

\item The orbital corresponding to the primary spin channel will be used to calculate DM elements of the fixed spin moments $\vec{S_i}$ and the secondary spin channel IAO will be used to calculate DM elements of the itinerant electrons $c_i^\dagger$
\end{enumerate}

\item \textbf{Optimally doped (x=0.125)}
\begin{enumerate}
\item The IAO analysis yields the same conclusions as the overdoped case.

\item Shown below is the difference in occupation of the spin-down and spin-up 3$d_{x^2-y^2}$ orbitals of the doped flip state MINUS the undoped flip state.

\includegraphics[width=0.5\textwidth]{../doped_fv/PBE0_8/FLP1/iao/sub_occd.pdf}
\linebreak

\includegraphics[width=0.5\textwidth]{../doped_fv/PBE0_8/FLP1/iao/DIFF_0_50.pdf} 
\linebreak

\includegraphics[width=0.5\textwidth]{../doped_fv/PBE0_8/FLP1/iao/DIFF_0_850.pdf}

\end{enumerate}
\end{enumerate}

\color{blue}
\subsection{Low energy states}
\begin{enumerate}
\item \textbf{Undoped (x=0)} For the undoped calculations, we have seven different spin textures. We can just use these for the model fitting since we will be fitting to a single parameter, $J$.

@@

\item \textbf{(Near) optimally doped (x=0.125)}
\begin{enumerate}
\item Same as x=0.25.
\end{enumerate}
\end{enumerate}
